### 2024-06-26 parts/prelim.tex
up to line 152

@@

\begin{example}
    \label{non-canonical-transd:ex}
    Let $\BadExOk \colon \set{a}^* \to \Nat$
    be defined as
    $\BadExOk(\varepsilon) = 1$,
    $\BadExOk(a) = 0$,
    and $\BadExOk(a^n) = n - 1$ for all $n \geq 2$.
    The function $\BadExOk$ is \kl{computed} by the two \kl{$\NPoly[0]$-transducers}
    depicted in \cref{non-canonical-transd:fig}.
    As a consequence, $f \in \NPoly[1]$.
    Furthermore, because the automaton of 
        \cref{non-canonical-transd:fig:counter-free}
        has no \kl{counter},
    and the correction function only uses functions in $\NSF[0]$,
    $f \in \NSF[1]$.
\end{example}

@@

            \draw[->] (B) to[bend right=45] node[midway, above] {$a \mid \ind{\card{w} > 1} + \ind{\card{w} \geq 1}$} (A);
        \end{tikzpicture}
        \caption{This automaton has a \kl{counter}.}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}
        \begin{tikzpicture}[
            etat/.style={minimum size=2em}
            ]
            \node[etat,state,initial,
                accepting by arrow,
                accepting text={$1$},
                accepting where=below,
                ] (A) at (0,0) {$\varepsilon$};
            \node[etat,state,
                accepting by arrow,
                accepting text={$0$},
                accepting where=below,
                ] (B) at (2,0) {$a$};

@@

            \draw[->] (B) edge[loop right] node[midway, right]
                {$a \mid 1 + \ind{\card{w} = 0}$} (B);
        \end{tikzpicture}
        \caption{This automaton is \kl{counter-free}.}
        \label{non-canonical-transd:fig:counter-free}
    \end{subfigure}
    \caption{Two \kl{$\NPoly[0]$-transducers}
    \kl{computing} the function $\BadExOk$ of \cref{non-canonical-transd:ex}.}
    \label{non-canonical-transd:fig}
\end{figure}

>>>>

\begin{example}
    \label{simple-transd:ex}
    Let $\mathsf{Const}$ be the set of constant functions from $\Sigma^*$ to $\Rel$,
    and let us write $\const{n}$ for the constant function equal to $n \in \Rel$.
    In 
    \cref{simple-transd:fig}, we provide a graphical depiction of a
    \kl{$\mathsf{Const}$-transducer} that \kl{computes} the function
    $f \colon \set{a,b}^* \to \Rel$ defined by
    $f(\varepsilon) = 1$,
    $f(aw) = 3 \times \card{w}$,
    and $f(bw) = 3 \times \card{w} + 1$.
\end{example}

\begin{figure}
    \centering
    \begin{tikzpicture}[
        etat/.style={minimum size=2em}
        ]
        \node[etat,state,initial,
            accepting by arrow,
            accepting text={$1$},
            accepting where=below,
            ] (A) at (0,0) {$q_0$};
        \node[etat,state,
            accepting by arrow,
            accepting text={$0$},
            accepting where=below,
            ] (B) at (2.5,0) {$q_1$};

        \draw[->] (A) to[bend right=20] node[midway, below] {$a \mid \const{0}$} (B);
        \draw[->] (A) to[bend left=20] node[midway, above] {$b \mid \const{1}$} (B);

        \draw[->] (B) edge[loop right] node[midway, right]
            {$a,b \mid \const{3}$} (B);
    \end{tikzpicture}
    \caption{A \kl{$\mathsf{Const}$-transducer}, where $\mathsf{Const}$ is the set of constant functions
        from $\set{a,b}^*$ to $\Rel$,
        with a set of states $Q \defined \set{ q_0, q_1 }$,
        an initial state $q_0$,
        an output function $F \colon Q \to \Rel$ such that $F(q_0) = 1$ and $F(q_1) = 0$,
        and correction functions $\lambda(q,a) = \const{0}$, $\lambda(q,b) = \const{1}$,
        and $\lambda(q_1,a) = \lambda(q_1,b) = \const{3}$.
    }
    \label{simple-transd:fig}
\end{figure}

@@

\begin{example}
    \label{non-canonical-transd:ex}
    Let $\BadExOk \colon \set{a}^* \to \Nat$
    be defined as
    $\BadExOk(\varepsilon) = 1$,
    $\BadExOk(a) = 0$,
    and $\BadExOk(a^n) = n-1$ for all $n \geq 2$.
    The function $\BadExOk$ is \kl{computed} by the two \kl{$\NPoly[0]$-transducers}
    depicted in \cref{non-canonical-transd:fig}.
    As a consequence, $f \in \NPoly[1]$.
    Furthermore, because the automaton of 
        \cref{non-canonical-transd:fig:counter-free}
        has no \kl{counter},
    and the correction function only uses functions in $\NSF[0]$,
    $f \in \NSF[1]$.
\end{example}
\begin{proofof}[non-canonical-transd:ex]
    Let us prove that the transducer of
    \cref{non-canonical-transd:fig:counter-free}
    computes $f$.
    We prove by induction on $w$
    that $A(q_0, w) = f(w)$,
    and that $A(q_1, w) = f(aw)$.
    Let us first remark that $f(au) = \card{u}$.

    When $w = \varepsilon$, $A(q_0, \varepsilon) = F(\varepsilon) = 1 = f(\varepsilon)$.
    Similarly, $A(q_1, \varepsilon) = F(q_1) = 0 = f(a)$.

    Assume that $w = au$. Then:
    \begin{align*}
        A(q_0, w) = A(q_0, au) &= \lambda(q_0, a)(u) + A(q_1, u) \\ 
                               &= 0 + A(q_1,u) \\
                               &= f(au) & \text{by induction hypothesis} \\
                               &= f(w) 
    \end{align*}
    Similarly,
    \begin{align*}
        A(q_1, w) = A(q_1, au) &= \lambda(q_1, a)(u) + A(q_1, u) \\ 
                               &= 1 + \ind{\card{u} = 0} + A(q_1,u) \\
                               &= 1 + \ind{\card{u} = 0} + f(au) & \text{by induction hypothesis} \\
                               &= \begin{cases}
                                    2 + f(a) & \text{if $u = \varepsilon$} \\
                                    1 + f(au) & \text{otherwise}
                               \end{cases}
                               \\
                               &= \begin{cases}
                                    2 & \text{if $u = \varepsilon$} \\
                                    1 + \card{u} & \text{otherwise}
                               \end{cases}
                               \\
                               &= \begin{cases}
                                   f(aa) & \text{if $u = \varepsilon$} \\
                                   f(
                               \end{cases}
    \end{align*}
\end{proofof}

@@

            \draw[->] (B) to[bend right=45] node[midway, above] {$a \mid 1 + \ind{\card{w} \geq 1}$} (A);
        \end{tikzpicture}
        \caption{This automaton has a \kl{counter}.}
    \end{subfigure}
    \begin{subfigure}[b]{0.49\linewidth}
        \begin{tikzpicture}[
            etat/.style={minimum size=2em}
            ]
            \node[etat,state,initial,
                accepting by arrow,
                accepting text={$1$},
                accepting where=below,
                ] (A) at (0,0) {$\varepsilon$};
            \node[etat,state,
                accepting by arrow,
                accepting text={$0$},
                accepting where=below,
                ] (B) at (2,0) {$a$};

@@

            \draw[->] (B) edge[loop right] node[midway, right]
                {$a \mid 1$} (B);
        \end{tikzpicture}
        \caption{This automaton is \kl{counter-free}.}
        \label{non-canonical-transd:fig:counter-free}
    \end{subfigure}
    \caption{Two \kl{$\NPoly[0]$-transducers}
    \kl{computing} the function $\BadExOk$ of \cref{non-canonical-transd:ex},
    that maps $a^n$ to $n$ for all $n \geq 2$, $a$ to $0$, and $\varepsilon$ to $1$.
    }
    \label{non-canonical-transd:fig}
\end{figure}
